Disentangling the spinal mechanisms of illusory heat and burning sensations in the Thermal Grill Illusion

### 2023-07-12 Manuscript.Rmd
pdf_and_word

  - id : "1"
    institution : "Center of Functionally Integrative Neuroscience (CFIN), Department of Clinical Medicine, Aarhus University"
  - id : "2"
    institution : "Interacting Minds Center (IMC), Department XXX, Aarhus University"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document:
    latex_engine: xelatex
  word_document:
bibliography: knitting files/Citations.bib
link-citations: yes
linkcolor: blue
always_allow_html: yes
---

@@

# Results and discussion
The Thermal Grill Illusion (TGI) is characterised by two key phenomena: thermosensory enhancement and illusory pain. Thermosensory enhancement refers to an amplified perception of heat or cold when cold and warm stimuli are simultaneously applied, as opposed to when each stimulus is presented individually or paired with a neutral temperature. Notably, the majority of individuals experience an intensification of heat rather than cold [@fardo_beyond_2020]. Illusory pain, on the other hand, denotes the perception of a burning sensation elicited by the pairing of warm and cold stimuli, an experience that is largely absent or significantly diminished when each stimulus is presented alone or combined with a neutral temperature. Thus, indicators of a stronger TGI are heightened warm and burning ratings, coupled with reduced cold ratings. 
To investigate thermosensory and burning components of the TGI, participants received pairs of temperatures on their forearms and were asked to quantify the levels of cold, warmth, and burning they experienced during each stimulation. These stimuli consisted of either cold-warm pairs (TGI stimuli), which potentially evoked an illusion of heat and pain, or control stimuli that involved pairing a cold or a warm stimulus of identical temperature as in the TGI condition with a baseline temperature of 30°C (non-TGI stimuli). All stimulation pairs were presented at a fixed distance on the skin, either within the same dermatome or across dermatomes that mapped onto non-adjacent spinal segments. In doing so, the stimuli elicited cold-related and warm-related neural activity that was either more focal or more widespread along the spinal cord. For each stimulation, participants reported their ratings using three sequential VAS scales ranging from 0, indicating the lack of a sensation, to 100, indicating an extreme sensation. For each instance of stimulation, participants were guided to focus their reporting on the sensations originating from a specific thermode. While the participants were unaware, this location corresponded to either the colder stimulus (Exp. 1) or the warmer stimulus (Exp. 2) of the paired temperatures. VAS ratings were analysed using zero-inflated beta regressions.

@@

```{r, results="hide", fig.width = 7.2, fig.height = 5, warning = F, message = F, echo = F, fig.cap = "Figure 1"}
ggdraw() +
    draw_image(magick::image_read(here::here("figures","figure1.PNG")), scale = 1)
```

```{r, results="hide", fig.width = 7.2, fig.height = 5, warning = F, message = F, echo = F, fig.cap = "Figure 2"}
plot1()
```

@@

```{r, results="hide", fig.width = 7.2, fig.height = 7.2, warning = F, message = F, echo = F, fig.cap = "Figure 3"}
plot2()
```

@@

## Participants
The study entailed two separate experiments, collectively involving 80 healthy volunteers. The sample consisted of `r descript_exp1_gender %>% filter(Gender == "F") %>% .$number` females and `r descript_exp1_gender %>% filter(Gender == "M") %>% .$number` males, mean age = `r descript_exp1_age$mean_age` years old (SD = `r descript_exp1_age$sd_age`, range = `r descript_exp1_age$min_age` - `r descript_exp1_age$max_age`) in Experiment 1,

and `r descript_exp2_gender %>% filter(Gender == "F") %>% .$number` females and `r descript_exp2_gender %>% filter(Gender == "M") %>% .$number` males and `r descript_exp2_gender %>% filter(Gender != "M" & Gender != "F") %>% .$number` non-binary (Female at birth), mean age = `r descript_exp2_age$mean_age` years old (SD = `r descript_exp2_age$sd_age`, range = `r descript_exp2_age$min_age` - `r descript_exp2_age$max_age`), in Experiment 2. The research methodology complied with the principles set forth in the Declaration of Helsinki and received ethical approval from the Institutional Review Board (IRB) at the Danish Neuroscience Center, Aarhus University, Denmark. Prior to commencing the study, all participants were fully informed about the procedures and provided their voluntary consent.

@@

>>>>

  - id : "1"
    institution : "Center of Functionally Integrative Neuroscience (CFIN), Department of Clinical Medicine, Aarhus University"
  - id : "2"
    institution : "Interacting Minds Center (IMC), Department XXX, Aarhus University"
date: "`r Sys.Date()`"
output:
  word_document:
  pdf_document:
    latex_engine: xelatex
  html_document: default
bibliography: knitting files/Citations.bib
link-citations: yes
linkcolor: blue
always_allow_html: yes
---

@@

\newpage
# Results and discussion
The Thermal Grill Illusion (TGI) is characterised by two key phenomena: thermosensory enhancement and illusory pain. Thermosensory enhancement refers to an amplified perception of heat or cold when cold and warm stimuli are simultaneously applied, as opposed to when each stimulus is presented individually or paired with a neutral temperature. Notably, the majority of individuals experience an intensification of heat rather than cold [@fardo_beyond_2020]. Illusory pain, on the other hand, denotes the perception of a burning sensation elicited by the pairing of warm and cold stimuli, an experience that is largely absent or significantly diminished when each stimulus is presented alone or combined with a neutral temperature. Thus, indicators of a stronger TGI are heightened warm and burning ratings, coupled with reduced cold ratings. 
To investigate thermosensory and burning components of the TGI, participants received pairs of temperatures on their forearms and were asked to quantify the levels of cold, warmth, and burning they experienced during each stimulation. These stimuli consisted of either cold-warm pairs (TGI stimuli), which potentially evoked an illusion of heat and pain, or control stimuli that involved pairing a cold or a warm stimulus of identical temperature as in the TGI condition with a baseline temperature of 30°C (non-TGI stimuli). All stimulation pairs were presented at a fixed distance on the skin, either within the same dermatome or across dermatomes that mapped onto non-adjacent spinal segments. In doing so, the stimuli elicited cold-related and warm-related neural activity that was either more focal or more widespread along the spinal cord. For each stimulation, participants reported their ratings using three sequential VAS scales ranging from 0, indicating the lack of a sensation, to 100, indicating an extreme sensation. For each instance of stimulation, participants were guided to focus their reporting on the sensations originating from a specific thermode. While the participants were unaware, this location corresponded to either the colder stimulus (Exp. 1) or the warmer stimulus (Exp. 2) of the paired temperatures. VAS ratings were analysed using zero-inflated beta regressions.

@@

```{r, results="hide", fig.width = 7.2, fig.height = 5, warning = F, message = F, echo = F, fig.cap = "Four experimental conditions depicting an exemplar cold and warm thermode placement within and across dermatomes. Within dermatomes, the relative location of the cold thermode was proximal or distal. Across dermatomes, the relative location of the corresponding spinal segments was rostral or caudal. "}
ggdraw() +
    draw_image(magick::image_read(here::here("figures","figure1.PNG")), scale = 1)
```

```{r, results="hide", fig.width = 7.2, fig.height = 5, warning = F, message = F, echo = F, fig.cap = " Difference in VAS ratings between within and across dermatome conditions for each type of stimulation (Non-TGI and TGI) in each VAS rating quality (cold, warm and burning). Positive values represent higher ratings within dermatomes, negative values represent higher ratings across dermatomes. Small dots are individual means, large dots are population means for each condition and error bars are 95% confidence intervals."}
plot1()
```

@@

```{r, results="hide", fig.width = 7.2, fig.height = 7.2, warning = F, message = F, echo = F, fig.cap = "Difference in VAS ratings between the two dermatome conditions. The top panel shows the difference between proximal and distal cold afferent locations (within dermatomes) and the bottom panel shows the difference between rostral and caudal cold afferent locations, relative to warm, across stimulation type (Non-TGI, TGI) and VAS rating quality (cold, warm and burning). Small points show data from each participant, large dots are means across trials for each condition, and error bars show 95% confidence intervals."}
plot2()
```

@@

## Participants
The study entailed two separate experiments, collectively involving 80 healthy volunteers. The sample consisted of `r descript_exp1_gender %>% filter(Gender == "F") %>% .$number` females and `r descript_exp1_gender %>% filter(Gender == "M") %>% .$number` males, mean age = `r descript_exp1_age$mean_age` years old (SD = `r descript_exp1_age$sd_age`, range = `r descript_exp1_age$min_age` - `r descript_exp1_age$max_age`) in Experiment 1, and `r descript_exp2_gender %>% filter(Gender == "F") %>% .$number` females and `r descript_exp2_gender %>% filter(Gender == "M") %>% .$number` males and `r descript_exp2_gender %>% filter(Gender != "M" & Gender != "F") %>% .$number` non-binary (Female at birth), mean age = `r descript_exp2_age$mean_age` years old (SD = `r descript_exp2_age$sd_age`, range = `r descript_exp2_age$min_age` - `r descript_exp2_age$max_age`), in Experiment 2. The research methodology complied with the principles set forth in the Declaration of Helsinki and received ethical approval from the Institutional Review Board (IRB) at the Danish Neuroscience Center, Aarhus University, Denmark. Prior to commencing the study, all participants were fully informed about the procedures and provided their voluntary consent.

@@

# References

::: {#refs}
:::
